---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
p_load(ggplot2, rethinking)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]? 

- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
-( Then implement a quadratic approximation (hint check paragraph 2.4.2!).)
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)
  

```{r}
#What's Riccardo's estimated knowledge of CogSci?
#First we define prior and likelihood

#prior
dens = 20
p_grid = seq(from = 0, to = 1, length.out = dens)
prior = rep(1, dens)
#1 because we are agnostic and don't know anything. assume each point is equally likely.

#likelihood
R_likelihood = dbinom(3, size = 6, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
R_unstd.posterior = R_likelihood * prior

#standardize the posterior so it sums to 1
R_posterior = R_unstd.posterior/sum(R_unstd.posterior)

#draw the plot
R_Data=data.frame(grid=p_grid,posterior=R_posterior,prior=prior,likelihood=R_likelihood)
ggplot(R_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") 

#What is the probability he knows more than chance?
sum(R_posterior[p_grid > 0.5])


#quadratic approximation
#to use map, provide a formula, a list of data, and a list of start values for the parameters
R_globe.qa = map(
  alist(
    c ~ dbinom(6, p) , #binomial likelihood
    p ~ dunif(0, 1)    #uniform prior
  ) ,
data = list(c = 3)) #count of corrects

#display summary of quadratic approximation
precis(R_globe.qa)

```


2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.

```{r}
#What's Kristian's estimated knowledge of CogSci?
#First we define prior and likelihood

#prior
dens = 20
p_grid = seq(from = 0, to = 1, length.out = dens)
prior = rep(1, dens)
#1 because we are agnostic and don't know anything. assume each point is equally likely.

#likelihood
K_likelihood = dbinom(2, size = 2, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
K_unstd.posterior = K_likelihood * prior

#standardize the posterior so it sums to 1
K_posterior = K_unstd.posterior/sum(K_unstd.posterior)

#draw the plot
K_Data=data.frame(grid=p_grid,posterior=K_posterior,prior=prior,likelihood=K_likelihood)
ggplot(K_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") 

```

```{r}
#What's Josh's estimated knowledge of CogSci?
#First we define prior and likelihood

#prior
dens = 20
p_grid = seq(from = 0, to = 1, length.out = dens)
prior = rep(1, dens)
#1 because we are agnostic and don't know anything. assume each point is equally likely.

#likelihood
J_likelihood = dbinom(160, size = 198, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
J_unstd.posterior = J_likelihood * prior

#standardize the posterior so it sums to 1
J_posterior = J_unstd.posterior/sum(J_unstd.posterior)

#draw the plot
J_Data=data.frame(grid=p_grid,posterior=J_posterior,prior=prior,likelihood=J_likelihood)
ggplot(J_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") 
```

```{r}
#What's Mikkel's estimated knowledge of CogSci?
#First we define prior and likelihood

#prior
dens = 20
p_grid = seq(from = 0, to = 1, length.out = dens)
prior = rep(1, dens)
#1 because we are agnostic and don't know anything. assume each point is equally likely.

#likelihood
M_likelihood = dbinom(66, size = 132, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
M_unstd.posterior = M_likelihood * prior

#standardize the posterior so it sums to 1
M_posterior = M_unstd.posterior/sum(M_unstd.posterior)

#draw the plot
M_Data=data.frame(grid=p_grid,posterior=M_posterior,prior=prior,likelihood=M_likelihood)
ggplot(M_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") 
```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r}
#What's Riccardo's estimated knowledge of CogSci?
#First we define prior and likelihood

#prior
dens = 20
p_grid = seq(from = 0, to = 1, length.out = dens)
prior2 = dnorm(p_grid, mean = 0.8,sd = 0.2)


#likelihood
R_likelihood = dbinom(3, size = 6, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
R_unstd.posterior = R_likelihood * prior2

#standardize the posterior so it sums to 1
R_posterior = R_unstd.posterior/sum(R_unstd.posterior)

#draw the plot
R_Data=data.frame(grid=p_grid,posterior=R_posterior,prior=prior2,likelihood=R_likelihood)
ggplot(R_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") 

```


```{r}
#What's Kristian's estimated knowledge of CogSci?
#First we define prior and likelihood

#likelihood
K_likelihood = dbinom(2, size = 2, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
K_unstd.posterior = K_likelihood * prior2

#standardize the posterior so it sums to 1
K_posterior = K_unstd.posterior/sum(K_unstd.posterior)

#draw the plot
K_Data=data.frame(grid=p_grid,posterior=K_posterior,prior=prior2,likelihood=K_likelihood)
ggplot(K_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") 

```

```{r}
#What's Josh's estimated knowledge of CogSci?
#First we define prior and likelihood


#likelihood
J_likelihood = dbinom(160, size = 198, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
J_unstd.posterior = J_likelihood * prior2

#standardize the posterior so it sums to 1
J_posterior = J_unstd.posterior/sum(J_unstd.posterior)

#draw the plot
J_Data=data.frame(grid=p_grid,posterior=J_posterior,prior=prior2,likelihood=J_likelihood)
ggplot(J_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") 
```

```{r}
#What's Mikkel's estimated knowledge of CogSci?
#First we define prior and likelihood



#likelihood
M_likelihood = dbinom(66, size = 132, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
M_unstd.posterior = M_likelihood * prior2

#standardize the posterior so it sums to 1
M_posterior = M_unstd.posterior/sum(M_unstd.posterior)

#draw the plot
M_Data=data.frame(grid=p_grid,posterior=M_posterior,prior=prior2,likelihood=M_likelihood)
ggplot(M_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") 
```


4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}
#Riccardo uniform prior

#likelihood
R_likelihood = dbinom(300, size = 600, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
R_unstd.posterior = R_likelihood * prior

#standardize the posterior so it sums to 1
R_posterior = R_unstd.posterior/sum(R_unstd.posterior)

#draw the plot
R_Data=data.frame(grid=p_grid,posterior=R_posterior,prior=prior,likelihood=R_likelihood)
ggplot(R_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") + labs(title = "Uniform prior") 
```

```{r}
#Riccardo 
#likelihood
R_likelihood = dbinom(300, size = 600, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
R_unstd.posterior = R_likelihood * prior2

#standardize the posterior so it sums to 1
R_posterior = R_unstd.posterior/sum(R_unstd.posterior)

#draw the plot
R_Data=data.frame(grid=p_grid,posterior=R_posterior,prior=prior2,likelihood=R_likelihood)
ggplot(R_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") + labs(title = "Updated prior")  
```

```{r}
#Kristian uniform prior

#likelihood
K_likelihood = dbinom(200, size = 200, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
K_unstd.posterior = K_likelihood * prior

#standardize the posterior so it sums to 1
K_posterior = K_unstd.posterior/sum(K_unstd.posterior)

#draw the plot
K_Data=data.frame(grid=p_grid,posterior=K_posterior,prior=prior,likelihood=K_likelihood)
ggplot(K_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") + labs(title = "Uniform prior") 
```

```{r}
#Kristian 
#likelihood
K_likelihood = dbinom(200, size = 200, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
K_unstd.posterior = K_likelihood * prior2

#standardize the posterior so it sums to 1
K_posterior = K_unstd.posterior/sum(K_unstd.posterior)

#draw the plot
K_Data=data.frame(grid=p_grid,posterior=K_posterior,prior=prior2,likelihood=K_likelihood)
ggplot(K_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") + labs(title = "Updated prior") 
```

```{r}
#Josh uniform prior

#likelihood
J_likelihood = dbinom(16000, size = 19800, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
J_unstd.posterior = J_likelihood * prior

#standardize the posterior so it sums to 1
J_posterior = J_unstd.posterior/sum(J_unstd.posterior)

#draw the plot
J_Data=data.frame(grid=p_grid,posterior=J_posterior,prior=prior,likelihood=J_likelihood)
ggplot(J_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") + labs(title = "Uniform prior") 
```

```{r}
#Josh 
#likelihood
J_likelihood = dbinom(16000, size = 19800, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
J_unstd.posterior = J_likelihood * prior2

#standardize the posterior so it sums to 1
J_posterior = J_unstd.posterior/sum(J_unstd.posterior)

#draw the plot
J_Data=data.frame(grid=p_grid,posterior=J_posterior,prior=prior2,likelihood=J_likelihood)
ggplot(R_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") + labs(title = "Updated prior") 
```

```{r}
#Mikkel uniform prior

#likelihood
M_likelihood = dbinom(6600, size = 13200, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
M_unstd.posterior = M_likelihood * prior

#standardize the posterior so it sums to 1
M_posterior = M_unstd.posterior/sum(M_unstd.posterior)

#draw the plot
M_Data=data.frame(grid=p_grid,posterior=M_posterior,prior=prior,likelihood=M_likelihood)
ggplot(M_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") + labs(title = "Uniform prior") 
```

```{r}
#Mikkel 
#likelihood
M_likelihood = dbinom(6600, size = 13200, prob = p_grid)
#3 is how many 1's (correct answers). 6 is how big is your data (how many possible 1's. 

#then we calculate the posterior
M_unstd.posterior = M_likelihood * prior2

#standardize the posterior so it sums to 1
M_posterior = M_unstd.posterior/sum(M_unstd.posterior)

#draw the plot
M_Data=data.frame(grid=p_grid,posterior=M_posterior,prior=prior2,likelihood=M_likelihood)
ggplot(M_Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability") + labs(title = "Updated prior") 
```

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?
  Don't need to code it, just conceptual.

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



